Statistical Theory\\ Chess Dataset Analysis

### 2024-09-29 Article/Article.tex
updated article
Updated Article

\author{
   \IEEEauthorblockN{Dor Boker, Itamar Nakar}
   \IEEEauthorblockA{
      I.D: , 325829000\\
      Email: , itamar.nakar@gmail.com
   }
}

@@

\section{Results}

>>>>

\author{
   \IEEEauthorblockN{Dor Boker, Itamar Nakar}
   \IEEEauthorblockA{
      I.D: 209271279 , 325829000\\
      Email: dorboker@gmail.com, itamar.nakar@gmail.com
   }
}

@@

\section{Results}
\subsection{Data Transformation}

To optimize the quality of our analysis, we undertook a cleaning process to remove entries that could distort or obscure meaningful insights from the dataset. Below are the key transformations:

\begin{itemize}
    \item \textbf{Uncompetitive Games}: Games marked as casual or non-competitive were excluded. Such games may not reflect players' full effort or strategic depth, and are thus not representative of a standard chess match.

    \item \textbf{Resignations}: Games concluded by resignation were removed. A resignation introduces a psychological factor into game length and outcome, which can skew the statistical distribution. By removing these games, we aim to focus on games decided by normal play.

    \item \textbf{Elo Adjustments}: New players are initially assigned an Elo rating of 1500, which can skew the distribution of player strength. To mitigate this, each player's rating was updated to reflect their most recent Elo score within the dataset.
\end{itemize}

After these trasformations, we are left with 5,827 games which we will now begin to analyze.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{gamma_fit.png}
    \caption{Gamma distribution fitted to the 'Turns' data}
    \label{fig:gamma_fit}
\end{figure}
\subsection{Distribution of 'Turns'}

To better understand the distribution of the 'Turns' variable, we attempted to fit it to several probability distribution families. The Gamma distribution was selected first due to its ability to model skewed data. However, as shown in Figure \ref{fig:gamma_fit}, the Gamma distribution is too broad for this dataset, suggesting it does not fully capture the underlying structure of 'Turns.'

To refine the fit, we combined the Gamma distribution with a Poisson distribution, which has a narrower shape. This combination results in a more accurate fit to the empirical data, as shown in Figure \ref{fig:gam_poi_fit}. The Gamma-Poisson mixture balances the excess spread of the Gamma distribution with the narrowness of the Poisson, providing a better representation of the 'Turns' distribution.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{gam_poi_fit.png}
    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
    \label{fig:gam_poi_fit}
\end{figure}
